---
title: "**Cadre Harmonisé (CH) & Integrated Phase Classification (IPC) Estimations for West & Central Africa**"
output: 
  flexdashboard::flex_dashboard:
    source_code: "https://github.com/olanderb/CHIPCdataRBD/"
    social: "menu"
  
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(DT)
library(plotly)
library(readxl)
library(leaflet)
library(sf)
library(leaflet.extras2)

cadre_harmonise_caf_ipc_orig <- read_excel("C:\\RBD_CHIPC_dataprocessing\\data\\processed\\cadre_harmonise_caf_ipc.xlsx")
cadre_harmonise_caf_ipc <- cadre_harmonise_caf_ipc_orig %>% filter(exercise_year == 2021 & exercise_label == "Jan-May") %>% mutate_at(vars(population:phase35), replace_na, 0) 

#make tables
#by country
cadre_harmonise_caf_ipc_country <- cadre_harmonise_caf_ipc %>% group_by(Country = adm0_name, chtype) %>% summarise(population = sum(population), phase1 = sum(phase1), phase2 = sum(phase2), phase3 =sum(phase3), phase4 = sum(phase4), phase5 = sum(phase5), phase35 = sum(phase35)) %>% mutate_if(is.numeric, funs(. / 1000000)) %>% mutate_if(is.numeric, round, 2)
#total
cadre_harmonise_caf_ipc_total <- cadre_harmonise_caf_ipc %>% group_by(chtype) %>% 
  summarise(population = sum(population), phase1 = sum(phase1), phase2 = sum(phase2), phase3 = sum(phase3), phase4 = sum(phase4), phase5 = sum(phase5), phase35 = sum(phase35)) %>% mutate(Country = "Total") %>% mutate_if(is.numeric, funs(. / 1000000)) %>% mutate_if(is.numeric, round, 1)
#put country and total together
cadre_harmonise_caf_ipc_all <- bind_rows(cadre_harmonise_caf_ipc_country, cadre_harmonise_caf_ipc_total)
#seperate current from projected
cadre_harmonise_caf_ipc_current <- cadre_harmonise_caf_ipc_all %>% filter(chtype == "current") %>% select(-chtype)
cadre_harmonise_caf_ipc_projected <- cadre_harmonise_caf_ipc_all %>% filter(chtype == "projected") %>% select(-chtype)

#make graphs 
CH_colors = c("phase1" = "#c6ffc7", "phase2" = "#ffe718", "phase3" = "#e88400", "phase4" = "#e02d00", "phase5" = "#5e0803") 
#long
cadre_harmonise_caf_ipc_countrylong <- cadre_harmonise_caf_ipc_country %>% select(-population, -phase35) %>% pivot_longer(cols = phase1:phase5, names_to = "phase", "value")
cadre_harmonise_caf_ipc_currentgraph <- cadre_harmonise_caf_ipc_countrylong %>% filter(chtype == "current") %>% ggplot(aes(x = Country, y = value, fill = phase)) +geom_bar(stat = "identity") +theme_minimal() +theme(axis.text.x = element_text(angle = 90)) +scale_fill_manual(values=CH_colors) +xlab("") +ylab("") +
      theme(
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()
      ) 
cadre_harmonise_caf_ipc_projectedgraph <- cadre_harmonise_caf_ipc_countrylong %>% filter(chtype == "projected") %>% ggplot(aes(x = Country, y = value, fill = phase)) +geom_bar(stat = "identity") +theme_minimal() +theme(axis.text.x = element_text(angle = 90)) +scale_fill_manual(values=CH_colors) +xlab("") +ylab("")+
      theme(
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()
      ) 

#add CH shapefile
CH_curr <- read_sf("C:\\RBD_CHIPC_dataprocessing\\data\\geo\\finalized_CHgeofiles\\wca_CHIPC_mar2021_current.gpkg")
CH_proj <- read_sf("C:\\RBD_CHIPC_dataprocessing\\data\\geo\\finalized_CHgeofiles\\wca_CHIPC_mar2021_projected.gpkg")

#add boundarry of countries
wca_shp0 <- read_sf("C:\\RBD_CHIPC_dataprocessing\\data\\geo\\finalized_CHgeofiles\\wca_shp0all.gpkg")

#reassign
pal <-  colorFactor(palette = c("#c6ffc7", "#ffe718", "#e88400","#e02d00"), 
              levels = c("1", "2", "3", "4"))
#final phase
CHmapleaflet_current <- CH_curr %>% leaflet() %>% addPolygons(weight = 0.25, fillOpacity = 0.75,
                                                       color = ~pal(CH_curr$phase_class),
                                                       popup = paste("Country:", CH_curr$adm0_name, "<br>",
                                                                    "Admin1:", CH_curr$adm1_name, "<br>",
                                                                    "Admin2:", CH_curr$adm2_name, "<br>",
                                                                    "Phase:", CH_curr$currentmar2021_phase_class, "<br>",
                                                                    "Total Population:", CH_curr$population, "<br>",
                                                                    "Phase 1 Population:", CH_curr$phase1, "<br>",
                                                                    "Phase 2 Population:", CH_curr$phase2, "<br>",
                                                                    "Phase 3 Population:", CH_curr$phase3, "<br>",
                                                                    "Phase 4 Population:", CH_curr$phase4, "<br>",
                                                                    "Phase 5 Population:", CH_curr$phase5, "<br>",
                                                                    "Phase 3-5 Population:", CH_curr$phase35, "<br>"),
                                                       highlight = highlightOptions(weight = 3, color = "black", bringToFront = FALSE)) %>% 
                                                       addLegend("bottomleft", pal = pal, values = ~phase_class,
            title = "Phase",
            opacity = 4) %>% addPolylines(data=wca_shp0, color="#000000", weight=2.5) %>%
  addEasyprint() %>% setView(lat= 13, lng=3, zoom=4)  %>%
  addEasyButton(easyButton(
    icon="fa-globe", title="reset to original view",
    onClick=JS("function(btn, map){ map.setView([13, 3], 4);}")))

CHmapleaflet_projected <- CH_proj %>%  leaflet() %>% addPolygons(weight = 0.25, fillOpacity = 0.75,
                                                       color = ~pal(CH_proj$phase_class),
                                                       popup = paste("Country:", CH_proj$adm0_name, "<br>",
                                                                    "Admin1:", CH_proj$adm1_name, "<br>",
                                                                    "Admin2:", CH_proj$adm2_name, "<br>",
                                                                    "Phase:", CH_proj$phase_class, "<br>",
                                                                    "Total Population:", CH_proj$population, "<br>",
                                                                    "Phase 1 Population:", CH_proj$phase1, "<br>",
                                                                    "Phase 2 Population:", CH_proj$phase2, "<br>",
                                                                    "Phase 3 Population:", CH_proj$phase3, "<br>",
                                                                    "Phase 4 Population:", CH_proj$phase4, "<br>",
                                                                    "Phase 5 Population:", CH_proj$phase5, "<br>",
                                                                    "Phase 3-5 Population:", CH_proj$phase35, "<br>"),
                                                       highlight = highlightOptions(weight = 3, color = "black", bringToFront = F)) %>% 
                                                       addLegend("bottomleft", pal = pal, values = ~phase_class,
            title = "Phase",
            opacity = 4) %>% addPolylines(data=wca_shp0, color="#000000", weight=2.5) %>%
  addEasyprint() %>% setView(lat= 13, lng=3, zoom=4)  %>%
  addEasyButton(easyButton(
    icon="fa-globe", title="reset to original view",
    onClick=JS("function(btn, map){ map.setView([13, 3], 4);}")))
#food consumption
CHmapleaflet_current_FC <- CH_curr %>% filter(admin0Name != "Central African Republic") %>% leaflet() %>% addPolygons(weight = 0.25, fillOpacity = 0.75,
                                                       color = ~pal(CH_curr$foodconsumption_phase),
                                                       popup = paste("Country:", CH_curr$adm0_name, "<br>",
                                                                    "Admin1:", CH_curr$adm1_name, "<br>",
                                                                    "Admin2:", CH_curr$adm2_name, "<br>",
                                                                    "Phase:", CH_curr$currentmar2021_foodconsumption_phase, "<br>"),
                                                       highlight = highlightOptions(weight = 3, color = "black", bringToFront = FALSE)) %>% 
                                                       addLegend("bottomleft", pal = pal, values = ~foodconsumption_phase,
            title = "Phase",
            opacity = 4) %>% addPolylines(data=wca_shp0, color="#000000", weight=2.5) %>%
  addEasyprint() %>% setView(lat= 13, lng=3, zoom=4)  %>%
  addEasyButton(easyButton(
    icon="fa-globe", title="reset to original view",
    onClick=JS("function(btn, map){ map.setView([13, 3], 4);}")))

CHmapleaflet_projected_FC <- CH_proj %>% filter(admin0Name != "Central African Republic") %>% leaflet() %>% addPolygons(weight = 0.25, fillOpacity = 0.75,
                                                       color = ~pal(CH_proj$foodconsumption_phase),
                                                       popup = paste("Country:", CH_proj$adm0_name, "<br>",
                                                                    "Admin1:", CH_proj$adm1_name, "<br>",
                                                                    "Admin2:", CH_proj$adm2_name, "<br>",
                                                                    "Phase:", CH_proj$foodconsumption_phase, "<br>"),
                                                       highlight = highlightOptions(weight = 3, color = "black", bringToFront = FALSE)) %>% 
                                                       addLegend("bottomleft", pal = pal, values = ~foodconsumption_phase,
            title = "Phase",
            opacity = 4) %>% addPolylines(data=wca_shp0, color="#000000", weight=2.5) %>%
  addEasyprint() %>% setView(lat= 13, lng=3, zoom=4)  %>%
  addEasyButton(easyButton(
    icon="fa-globe", title="reset to original view",
    onClick=JS("function(btn, map){ map.setView([13, 3], 4);}")))
#livelihoods
CHmapleaflet_current_LH <- CH_curr %>% filter(admin0Name != "Central African Republic") %>% leaflet() %>% addPolygons(weight = 0.25, fillOpacity = 0.75,
                                                       color = ~pal(CH_curr$livelihoods_phase),
                                                       popup = paste("Country:", CH_curr$adm0_name, "<br>",
                                                                    "Admin1:", CH_curr$adm1_name, "<br>",
                                                                    "Admin2:", CH_curr$adm2_name, "<br>",
                                                                    "Phase:", CH_curr$livelihoods_phase, "<br>"),
                                                       highlight = highlightOptions(weight = 3, color = "black", bringToFront = FALSE)) %>% 
                                                       addLegend("bottomleft", pal = pal, values = ~livelihoods_phase,
            title = "Phase",
            opacity = 4) %>% addPolylines(data=wca_shp0, color="#000000", weight=2.5) %>%
  addEasyprint() %>% setView(lat= 13, lng=3, zoom=4)  %>%
  addEasyButton(easyButton(
    icon="fa-globe", title="reset to original view",
    onClick=JS("function(btn, map){ map.setView([13, 3], 4);}")))

CHmapleaflet_projected_LH <- CH_proj %>% filter(admin0Name != "Central African Republic") %>% leaflet() %>% addPolygons(weight = 0.25, fillOpacity = 0.75,
                                                       color = ~pal(CH_proj$livelihoods_phase),
                                                       popup = paste("Country:", CH_proj$adm0_name, "<br>",
                                                                    "Admin1:", CH_proj$adm1_name, "<br>",
                                                                    "Admin2:", CH_proj$adm2_name, "<br>",
                                                                    "Phase:", CH_proj$livelihoods_phase, "<br>"),
                                                       highlight = highlightOptions(weight = 3, color = "black", bringToFront = FALSE)) %>% 
                                                       addLegend("bottomleft", pal = pal, values = ~livelihoods_phase,
            title = "Phase",
            opacity = 4) %>% addPolylines(data=wca_shp0, color="#000000", weight=2.5) %>%
  addEasyprint() %>% setView(lat= 13, lng=3, zoom=4)  %>%
  addEasyButton(easyButton(
    icon="fa-globe", title="reset to original view",
    onClick=JS("function(btn, map){ map.setView([13, 3], 4);}")))

#nutrition
CHmapleaflet_current_Nut <- CH_curr %>% filter(admin0Name != "Central African Republic") %>% leaflet() %>% addPolygons(weight = 0.25, fillOpacity = 0.75,
                                                       color = ~pal(CH_curr$nutrition_phase),
                                                       popup = paste("Country:", CH_curr$adm0_name, "<br>",
                                                                    "Admin1:", CH_curr$adm1_name, "<br>",
                                                                    "Admin2:", CH_curr$adm2_name, "<br>",
                                                                    "Phase:", CH_curr$currentmar2021_nutrition_phase, "<br>"),
                                                       highlight = highlightOptions(weight = 3, color = "black", bringToFront = FALSE)) %>% 
                                                       addLegend("bottomleft", pal = pal, values = ~nutrition_phase,
            title = "Phase",
            opacity = 4) %>% addPolylines(data=wca_shp0, color="#000000", weight=2.5) %>%
  addEasyprint() %>% setView(lat= 13, lng=3, zoom=4)  %>%
  addEasyButton(easyButton(
    icon="fa-globe", title="reset to original view",
    onClick=JS("function(btn, map){ map.setView([13, 3], 4);}")))

CHmapleaflet_projected_Nut <- CH_proj %>% filter(admin0Name != "Central African Republic") %>% leaflet() %>% addPolygons(weight = 0.25, fillOpacity = 0.75,
                                                       color = ~pal(CH_proj$projectedmar2021_nutrition_phase),
                                                       popup = paste("Country:", CH_proj$adm0_name, "<br>",
                                                                    "Admin1:", CH_proj$adm1_name, "<br>",
                                                                    "Admin2:", CH_proj$adm2_name, "<br>",
                                                                    "Phase:", CH_proj$projectedmar2021_nutrition_phase, "<br>"),
                                                       highlight = highlightOptions(weight = 3, color = "black", bringToFront = FALSE)) %>% 
                                                       addLegend("bottomleft", pal = pal, values = ~nutrition_phase,
            title = "Phase",
            opacity = 4) %>% addPolylines(data=wca_shp0, color="#000000", weight=2.5) %>%
  addEasyprint() %>% setView(lat= 13, lng=3, zoom=4)  %>%
  addEasyButton(easyButton(
    icon="fa-globe", title="reset to original view",
    onClick=JS("function(btn, map){ map.setView([13, 3], 4);}")))

#mortality
CHmapleaflet_current_Mort <- CH_curr %>% filter(admin0Name != "Central African Republic") %>% leaflet() %>% addPolygons(weight = 0.25, fillOpacity = 0.75,
                                                       color = ~pal(CH_curr$mortality_phase),
                                                       popup = paste("Country:", CH_curr$adm0_name, "<br>",
                                                                    "Admin1:", CH_curr$adm1_name, "<br>",
                                                                    "Admin2:", CH_curr$adm2_name, "<br>",
                                                                    "Phase:", CH_curr$mortality_phase, "<br>"),
                                                       highlight = highlightOptions(weight = 3, color = "black", bringToFront = FALSE)) %>% 
                                                       addLegend("bottomleft", pal = pal, values = ~mortality_phase,
            title = "Phase",
            opacity = 4) %>% addPolylines(data=wca_shp0, color="#000000", weight=2.5) %>%
  addEasyprint() %>% setView(lat= 13, lng=3, zoom=4)  %>%
  addEasyButton(easyButton(
    icon="fa-globe", title="reset to original view",
    onClick=JS("function(btn, map){ map.setView([13, 3], 4);}")))

CHmapleaflet_projected_Mort <- CH_proj %>% filter(admin0Name != "Central African Republic") %>% leaflet() %>% addPolygons(weight = 0.25, fillOpacity = 0.75,
                                                       color = ~pal(CH_proj$mortality_phase),
                                                       popup = paste("Country:", CH_proj$adm0_name, "<br>",
                                                                    "Admin1:", CH_proj$adm1_name, "<br>",
                                                                    "Admin2:", CH_proj$adm2_name, "<br>",
                                                                    "Phase:", CH_proj$projectedmar2021_mortality_phase, "<br>"),
                                                       highlight = highlightOptions(weight = 3, color = "black", bringToFront = FALSE)) %>% 
                                                       addLegend("bottomleft", pal = pal, values = ~mortality_phase,
            title = "Phase",
            opacity = 4) %>% addPolylines(data=wca_shp0, color="#000000", weight=2.5) %>%
  addEasyprint() %>% setView(lat= 13, lng=3, zoom=4)  %>%
  addEasyButton(easyButton(
    icon="fa-globe", title="reset to original view",
    onClick=JS("function(btn, map){ map.setView([13, 3], 4);}")))




#for making trends since 2017
#only since 2017 and most relevant period 
cadre_harmonise_caf_since2017 <- cadre_harmonise_caf_ipc_orig %>% filter(exercise_year >= 2017)  %>% mutate_at(vars(population:phase35), replace_na, 0) %>% filter(usethisperiod == "Y")

#trends of phase35 in millions
#create tables for graphing
cadre_harmonise_caf_since2017_ph35 <- cadre_harmonise_caf_since2017 %>% group_by(adm0_name, reference_label, reference_year) %>% summarise(phase35 = sum(phase35)) %>% mutate(phase35 = round((phase35 / 1000000),1)) 
graph_sahel <- cadre_harmonise_caf_since2017_ph35 %>% filter(adm0_name %in% c(
  "Burkina Faso","Cameroon","Central African Republic", "Chad","Mali","Mauritania","Niger","Nigeria","Senegal")) %>% ggplot(aes(x = reference_year, y = phase35, group = adm0_name, color = adm0_name)) +geom_line() +facet_grid(. ~ reference_label) +theme_minimal() +theme(axis.title.x = element_blank()) +labs(color='Country') 
graph_coastal <- cadre_harmonise_caf_since2017_ph35 %>% filter(adm0_name %in% c(
  "Benin","Cabo Verde","Cote d'Ivoire","Gambia","Ghana","Guinea","Guinea-Bissau","Liberia","Sierra Leone","Togo")) %>% ggplot(aes(x = reference_year, y = phase35, group = adm0_name, color = adm0_name)) +geom_line() +facet_grid(. ~ reference_label) +theme_minimal() +theme(axis.title.x = element_blank()) +labs(color='Country') 
#create tables of % phase35 of analyzed population
cadre_harmonise_caf_since2017 <- cadre_harmonise_caf_since2017 %>% mutate(totalpop = phase1+phase2+phase3+phase4+phase5)
cadre_harmonise_caf_since2017perc <- cadre_harmonise_caf_since2017 %>% group_by(adm0_name, reference_label, reference_year) %>% summarise(phase35 = sum(phase35), totalpop = sum(totalpop)) %>% mutate(percphase35 = round((phase35 / totalpop) * 100),1)
graph_sahelperc <- cadre_harmonise_caf_since2017perc %>% filter(adm0_name %in% c(
  "Burkina Faso","Cameroon","Central African Republic", "Chad","Mali","Mauritania","Niger","Nigeria","Senegal")) %>% ggplot(aes(x = reference_year, y = percphase35, group = adm0_name, color = adm0_name)) +geom_line() +facet_grid(. ~ reference_label) +theme_minimal() +theme(axis.title.x = element_blank()) +labs(color='Country') 
graph_coastalperc <- cadre_harmonise_caf_since2017perc %>% filter(adm0_name %in% c(
  "Benin","Cabo Verde","Cote d'Ivoire","Gambia","Ghana","Guinea","Guinea-Bissau","Liberia","Sierra Leone","Togo")) %>% ggplot(aes(x = reference_year, y = percphase35, group = adm0_name, color = adm0_name)) +geom_line() +facet_grid(. ~ reference_label) +theme_minimal() +theme(axis.title.x = element_blank()) +labs(color='Country') 




```

Map of Current & Projected Phases {data-navmenu="Maps"}
=====================================

Row 
-------------------------------------

### Phasing of Areas: Current (January - May 2021)  

```{r results="asis"}
cat("
<style>
.leaflet-container {
    background: #FFF;
}
</style>
")
CHmapleaflet_current
```  

> Click on polygon area for more information

Row 
-------------------------------------

### Phasing of Areas:  Projected (June - August 2021)

```{r results="asis"}
cat("
<style>
.leaflet-container {
    background: #FFF;
}
</style>
")
CHmapleaflet_projected
```  


Population Estimates (table & graph) {data-navmenu="Figures"}
=====================================

Row 
-------------------------------------

### Population Estimates (in millions) by Phase: March 2021 Exercise (Current: Jan.-May. 2021)  

```{r}
ggplotly(cadre_harmonise_caf_ipc_currentgraph)
```  
  
> Hover cursor over graph for more information; select phases in legend to filter

### Population Estimates (in millions) by Phase: March 2021 Exercise (Current: Jan.- May. 2021)  

```{r}
datatable(cadre_harmonise_caf_ipc_current,  extensions = 'Buttons', options = list(pageLength = 18, dom = "Bft",
    buttons = c('copy', 'csv', 'excel'))) %>%
  formatStyle(
    0,
    target = "row",
    fontWeight = styleEqual(18, "bold")
  )

```

Row 
-------------------------------------


### Population Estimates (in millions) by Phase: March 2021 Exercise (Projected: Jun. - Aug. 2021) 

```{r}
ggplotly(cadre_harmonise_caf_ipc_projectedgraph)
``` 

> Hover cursor over graph for more information; select phases in legend to filter

### Population Estimates (in millions) by Phase: March 2021 Exercise (Projected: Jun. - Aug. 2021)
    
```{r}
datatable(cadre_harmonise_caf_ipc_projected,  extensions = 'Buttons', options = list(pageLength = 19, dom = "Bft",
    buttons = c('copy', 'csv', 'excel'))) %>%
  formatStyle(
    0,
    target = "row",
    fontWeight = styleEqual(19, "bold")
  )
```


Phase 3-5 Trends (population) {data-navmenu="Trends"}
=====================================

Row 
-------------------------------------

### Trends of the number of people in phas3 - 5 (in millions) by Reference Period: Sahel5 + Cameroon, C.A.R. and Nigeria

```{r}
ggplotly(graph_sahel)
``` 

> Hover cursor over graph for more information; select countries in legend to filter

### Trends of the number of people in phas3 - 5 (in millions) by Reference Period: Coastal Countries

```{r}
ggplotly(graph_coastal)
``` 

> Hover cursor over graph for more information; select countries in legend to filter

Phase 3-5 Trends (percent of population) {data-navmenu="Trends"}
=====================================

Row 
-------------------------------------

### Trends of the percent of people in phase3 - 5 out of total analyzed population by Reference Period: Sahel5 + Cameroon, C.A.R. and Nigeria

```{r}
ggplotly(graph_sahelperc)
``` 

> Hover cursor over graph for more information; select countries in legend to filter

### Trends of the percent of people in phase3 - 5 out of total analyzed population by Reference Period: Coastal Countries

```{r}
ggplotly(graph_coastalperc)
``` 

> Hover cursor over graph for more information; select countries in legend to filter