---
title: "**Cadre Harmonisé (CH) & Integrated Phase Classification (IPC) Estimations for West & Central Africa November 2020 Exercise Results**"
output: 
  flexdashboard::flex_dashboard:
  orientation: rows
  vertical_layout: scroll
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(DT)
library(plotly)
library(readxl)
library(leaflet)
library(sf)
library(leaflet.extras2)

cadre_harmonise_caf_ipc <- read_excel("cadre_harmonise_caf_ipc.xlsx")
cadre_harmonise_caf_ipcnov2020 <- cadre_harmonise_caf_ipc %>% filter(exercise_year == 2020 & exercise_label == "Sep-Dec") %>% mutate_at(vars(population:phase35), replace_na, 0) 

#make tables
#by country
cadre_harmonise_caf_ipcnov2020country <- cadre_harmonise_caf_ipcnov2020 %>% group_by(Country = adm0_name, chtype) %>% summarise(population = sum(population), phase1 = sum(phase1), phase2 = sum(phase2), phase3 =sum(phase3), phase4 = sum(phase4), phase5 = sum(phase5), phase35 = sum(phase35)) %>% mutate_if(is.numeric, funs(. / 1000000)) %>% mutate_if(is.numeric, round, 2)
#total
cadre_harmonise_caf_ipcnov2020total <- cadre_harmonise_caf_ipcnov2020 %>% group_by(chtype) %>% 
  summarise(population = sum(population), phase1 = sum(phase1), phase2 = sum(phase2), phase3 = sum(phase3), phase4 = sum(phase4), phase5 = sum(phase5), phase35 = sum(phase35)) %>% mutate(Country = "Total") %>% mutate_if(is.numeric, funs(. / 1000000)) %>% mutate_if(is.numeric, round, 1)
#put country and total together
cadre_harmonise_caf_ipcnov2020all <- bind_rows(cadre_harmonise_caf_ipcnov2020country, cadre_harmonise_caf_ipcnov2020total)
#seperate current from projected
cadre_harmonise_caf_ipcnov2020current <- cadre_harmonise_caf_ipcnov2020all %>% filter(chtype == "current") %>% select(-chtype)
cadre_harmonise_caf_ipcnov2020projected <- cadre_harmonise_caf_ipcnov2020all %>% filter(chtype == "projected") %>% select(-chtype)

#make graphs 
CH_colors = c("phase1" = "#c6ffc7", "phase2" = "#ffe718", "phase3" = "#e88400", "phase4" = "#e02d00", "phase5" = "#5e0803") 
#long
cadre_harmonise_caf_ipcnov2020countrylong <- cadre_harmonise_caf_ipcnov2020country %>% select(-population, -phase35) %>% pivot_longer(cols = phase1:phase5, names_to = "phase", "value")
cadre_harmonise_caf_ipcnov2020currentgraph <- cadre_harmonise_caf_ipcnov2020countrylong %>% filter(chtype == "current") %>% ggplot(aes(x = Country, y = value, fill = phase)) +geom_bar(stat = "identity") +theme_minimal() +theme(axis.text.x = element_text(angle = 90)) +scale_fill_manual(values=CH_colors) +xlab("") +ylab("") +
      theme(
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()
      ) 
cadre_harmonise_caf_ipcnov2020projectedgraph <- cadre_harmonise_caf_ipcnov2020countrylong %>% filter(chtype == "projected") %>% ggplot(aes(x = Country, y = value, fill = phase)) +geom_bar(stat = "identity") +theme_minimal() +theme(axis.text.x = element_text(angle = 90)) +scale_fill_manual(values=CH_colors) +xlab("") +ylab("")+
      theme(
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()
      ) 

#add CH shapefile
CH <- read_sf("wca_CHIPC_clean_nov2020.gpkg")
#add boundarry of countries
wca_shp0 <- read_sf("wca_shp0.gpkg")

#reassign
pal <-  colorFactor(palette = c("#c6ffc7", "#ffe718", "#e88400","#e02d00","#E1E1E1"), 
              levels = c("1", "2", "3", "4","Not Analyzed"))

CHmapleaflet_current <- CH %>% leaflet() %>% addPolygons(weight = 0.25, fillOpacity = 0.75,
                                                       color = ~pal(currentnov2020_phase_class),
                                                       popup = paste("Country:", CH$adm0_name, "<br>",
                                                                    "Admin1:", CH$adm1_name, "<br>",
                                                                    "Admin2:", CH$adm2_name, "<br>",
                                                                    "Phase:", CH$currentnov2020_phase_class, "<br>",
                                                                    "Total Population:", CH$currentnov2020_population, "<br>",
                                                                    "Phase 1 Population:", CH$currentnov2020_phase1, "<br>",
                                                                    "Phase 2 Population:", CH$currentnov2020_phase2, "<br>",
                                                                    "Phase 3 Population:", CH$currentnov2020_phase3, "<br>",
                                                                    "Phase 4 Population:", CH$currentnov2020_phase4, "<br>",
                                                                    "Phase 5 Population:", CH$currentnov2020_phase5, "<br>",
                                                                    "Phase 3-5 Population:", CH$currentnov2020_phase35, "<br>"),
                                                       highlight = highlightOptions(weight = 3, color = "black", bringToFront = FALSE)) %>% 
                                                       addLegend("bottomleft", pal = pal, values = ~currentnov2020_phase_class,
            title = "Phase",
            opacity = 4) %>% addPolylines(data=wca_shp0, color="#000000", weight=2.5) %>%
  addEasyprint() %>% setView(lat= 13, lng=3, zoom=4)  %>%
  addEasyButton(easyButton(
    icon="fa-globe", title="reset to original view",
    onClick=JS("function(btn, map){ map.setView([13, 3], 4);}")))




CHmapleaflet_projected <- CH %>% leaflet() %>% addPolygons(weight = 0.25, fillOpacity = 0.75,
                                                       color = ~pal(projectednov2020_phase_class),
                                                       popup = paste("Country:", CH$adm0_name, "<br>",
                                                                    "Admin1:", CH$adm1_name, "<br>",
                                                                    "Admin2:", CH$adm2_name, "<br>",
                                                                    "Phase:", CH$projectednov2020_phase_class, "<br>",
                                                                    "Total Population:", CH$projectednov2020_population, "<br>",
                                                                    "Phase 1 Population:", CH$projectednov2020_phase1, "<br>",
                                                                    "Phase 2 Population:", CH$projectednov2020_phase2, "<br>",
                                                                    "Phase 3 Population:", CH$projectednov2020_phase3, "<br>",
                                                                    "Phase 4 Population:", CH$projectednov2020_phase4, "<br>",
                                                                    "Phase 5 Population:", CH$projectednov2020_phase5, "<br>",
                                                                    "Phase 3-5 Population:", CH$currentnov2020_phase35, "<br>"),
                                                       highlight = highlightOptions(weight = 3, color = "black", bringToFront = F)) %>% 
                                                       addLegend("bottomleft", pal = pal, values = ~projectednov2020_phase_class,
            title = "Phase",
            opacity = 4) %>% addPolylines(data=wca_shp0, color="#000000", weight=2.5) %>%
  addEasyprint() %>% setView(lat= 13, lng=3, zoom=4)  %>%
  addEasyButton(easyButton(
    icon="fa-globe", title="reset to original view",
    onClick=JS("function(btn, map){ map.setView([13, 3], 4);}")))







```

Map of Current & Projected Phases {data-navmenu="Maps"}
=====================================

Row 
-------------------------------------

### Phasing of Areas: Current (September - December 2020)  
```{r, fig.height= 10, fig.width= 5}
CHmapleaflet_current
```  

> Click on polygon area for more information

Row 
-------------------------------------

### Phasing of Areas:  Projected (June - August 2021)

```{r, fig.height= 10, fig.width= 5}
CHmapleaflet_projected
```  

> Click on polygon area for more information


Population Estimates (table & graph) {data-navmenu="Figures"}
=====================================

Row 
-------------------------------------

### Population Estimates (in millions) by Phase: November 2020 Exercise Current (September - December 2020)  

```{r}
ggplotly(cadre_harmonise_caf_ipcnov2020currentgraph)
```  
  
> Hover cursor over graph for more information; select phases in legend to filter

### Population Estimates (in millions) by Phase: November 2020 Exercise Current (September - December 2020)  

```{r}
datatable(cadre_harmonise_caf_ipcnov2020current,  extensions = 'Buttons', options = list(pageLength = 18, dom = "Bft",
    buttons = c('copy', 'csv', 'excel'))) %>%
  formatStyle(
    0,
    target = "row",
    fontWeight = styleEqual(18, "bold")
  )

```

Row 
-------------------------------------


### Population Estimates (in millions) by Phase: November 2020 Exercise Projected (June - August 2021) 

```{r}
ggplotly(cadre_harmonise_caf_ipcnov2020projectedgraph)
``` 

> Hover cursor over graph for more information; select phases in legend to filter

### Population Estimates (in millions) by Phase: November 2020 Exercise Projected (June - August 2021)
    
```{r}
datatable(cadre_harmonise_caf_ipcnov2020projected,  extensions = 'Buttons', options = list(pageLength = 18, dom = "Bft",
    buttons = c('copy', 'csv', 'excel'))) %>%
  formatStyle(
    0,
    target = "row",
    fontWeight = styleEqual(18, "bold")
  )
```